Sketch Recognition and Understanding

### 2009-07-28 draft/chapterSurvey.tex

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize independent symbols. Each symbol is segmented and hierarchical features are computed.  Features that are representing the global description of the shape are also generated and appended to the hierarchical features. Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

\chapter{Particle Swarm Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swarm Optimization Review}
\label{sec:ParticleSwramOptimizationReview}
%This seems to emerge of what is called 'collective intelligence' from often large simple agents.  
% start with swarm in general  .. say about flock , ant and usin in image processing 
% then speeak about PSO and  its firstly generated   (see article)
%mention that is best than ga (see thesis)

Modeling how a biological swarm has been recently studied by biologist and computer scientists. Scientists in the field of artificial life try to understand how such "social animals" interact, achieve goals and evolve. Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as they can apply it in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal. A good example of this cooperation is ant groups looking for food source. A large group of simple independent agents seems to generate this behavior in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

@@

Image processing is one of the disciplines that used Swarm Intelligence. Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}. Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}. PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process.

\section{Discrete Particle Swarm Algorithm}
\label{sec:ParticleSwarmAlgorithm}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each agent with a particle from the solution space \cite{PSOFirst}. Each agent moves the particle with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm} \& \ref{eq:Swarm1}.
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij}  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
\label{eq:Swarm}
\end{equation}
 where $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm system variables.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}. 
 \begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i} 
\label{eq:descrite}
\end{array}\}
\end{equation}
 where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable.

>>>>

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize independent symbols. Each symbol is segmented and hierarchical features are computed.  Features that are representing the global description of the shape are also generated and appended to the hierarchical features. Recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms. 
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
%*********************************************************************************************5&&&&&&&&&&&&&&&&&&&&
\chapter{Particle Swarm Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swarm Intelligence Review}
\label{sec:ParticleSwramOptimizationReview}

 Modeling how a biological swarm has been recently studied by biologist and computer scientists. Scientists in the field of artificial life try to understand how such "social animals" interact, achieve goals and evolve. Moreover, a swarm behavior that achieves some kind of swarm intelligence is increasingly interesting for engineers as they can apply it in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal. A good example of this cooperation is ant groups looking for food source. A large group of simple independent agents seems to generate this behavior in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}. %with other neighborhood agents.

@@

The underling research in swarm intelligence led to optimization algorithm that simulates the animal social behaviors. Ant Colony Optimization (ACO) algorithm simulate how ants search for optimize path for food \cite{ACOFirst}. Real ants leave evidence of their exploration while they are searching for food. This trail is used by other ants to locate shortest distance between the colony and food location. ACO use similar type of cooperation between agents to find optimal solution.  The synchronous bird flock movement was simulated in \cite{PSOFirst,PSO2}.  Kennedy and Eberhart \cite{PSOFirst,PSO2} simulated the rules of bird flock movement and introduced the Particle Swarm Optimization Algorithm (PSO).

   The next section describes in details the Particle Swarm Optimization and its advantages and disadvantages. 
% Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO). The algorithm simulate the moving Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) as a general optimization algorithm.  
 \subsection{Particle Swarm Optimization}
 In optimization problems, the algorithm attempts to find a feasible solution which satisfies all constrain of this problem. An optimal solution is the best feasible solution in the solution space. Several methods were used to solve optimization problems. Stochastic search is based on the assumption that good solution are near each other. Algorithms like simulating annealing and hill climbing used stochastic to find best solution. The problem in these methods is they may be trapped into a local best and fail to find the global best solution.

  Evolutionary algorithms (e.g. Genatic Algoirthms) explore solution using a population which is evaluated using fitness functions. The algorithms initially start with a randomly chosen population which is used to generate new population using various operators (e.g. cross over, mutation... ).  After each iteration, the population is evaluated using a fitness function and only the strongest are used to generate new population. The use of population helps to avoid stochastic search problem and explores the solution space to find better global optimal solutions, but unfortunately they may not always converge to a global best solution\cite{PSOpattern}.

 Particle Swarm Algorithm\cite{PSOFirst}is a population based stochastic search where each solution is simulated as a  particle that flies around in a multidimensional search space. A fitness function is used to evaluate the goodness of each particle in solution space. But unlike evolutionally algorithm the new population is the PSO is influenced by the social cooperation of the particle rather than survival to the fittest. Each particle saves a history of the best location in the swarm and its own best location.  At the end each iteration, the particles moves to a new location based on the global best location and it own history of best location. This use of pervious particle locations gives PSO advantage over Evolutionary algorithms where no similar information is used.

\subsection{PSO vs. GA}
 PSO can be viewed as an algorithm between stochastic search and Evolutionary algorithms\cite{PSOpattern}. In general the PSO is a continuous optimization problem but experiments showed it performed as well as GA which is in general a discrete and binary optimization problem. In general, both GA and PSO share many points, e.g, both algorithms generate the initial populuation randomaly, both use a fitness function or value to evaluate the goodness of solutions and the next populations. Nevertheless, PSO dose not have genetic operators like crossover or mutations. Partices are updated by internal comunication among themselfs based on the local and global best. In contrast to GA where chromosomes shares information with each others, PSO shares information using memory from previous iteration in form of global best solution. This leads the population to move as one group to towards an optimal solution (i.e find solution quickly). %which significally

  %From the procedure, we can learn that PSO shares many common points with GA. Both algorithms start with a group of a randomly generated population, both have fitness values to evaluate the population. Both update the population and search for the optimium with random techniques. Both systems do not guarantee success.

%However, PSO does not have genetic operators like crossover and mutation. Particles update themselves with the internal velocity. They also have memory, which is important to the algorithm.

%Compared with genetic algorithms (GAs), the information sharing mechanism in PSO is significantly different. In GAs, chromosomes share information with each other. So the whole population moves like a one group towards an optimal area. In PSO, only gBest (or lBest) gives out the information to others. It is a one -way information sharing mechanism. The evolution only looks for the best solution. Compared with GA, all the particles tend to converge to the best solution quickly even in the local version in most cases.
 The comparison between PSO and Genetic algorithm was performed in \cite{PSOGAComp,ComparePSOGA05,CoPSAanGA05,PAOvsGA07} and proved that PSO generate faster, better optimal solutions than GA especially in problems with high dimensionality. In addition, Rohinsons et al. \cite{RohinsonPSGA} proved that PSO performed better than GA in difficult engineering problems \cite{EngineerProblemGA}.

% Eberhart \cite{PSOGAComp}
%%PSO has been shown to have the same ability to find a global optimum as genetic
%algorithms, but is able to find optimums faster than genetic algorithms (Panda and Padhy
%2007; Mouser and Dunn 2005; Hassan et al. 2005)

 A hybrid of PSO and other algorithms was investigated in various systems. Veeramachaneni \cite{kaylanGAPSO} used a hybrid of GA and PSO to breed solutions using GA then feed them to PSO. They demonstrated that using PSO-GA and PSO outperform using GA as an optimization algorithm.

\subsection{PSO Neighborhood Topology}  
In general the PSO algorithm have fully connected particle which means that every particle know the global best and local best of each other particle. Other forms of cooperation and organization were suggested in various systems.  \cite{grid} suggested using a grid like topology where particle are connected to four other particles around it. A ring and star topology were widely used in \cite{neighbourrhood}. The advantages over using different topology than the global best is that they slow down the swarm evolving (they need more time to propagate the best solution) which enable more space exploration and thus helps in eliminating the premature convergence problem. In \cite{hiericalPSO}, Janson and Middendorf proposed to use a dynamic and hierarchical neighborhood structure. They showed that using such structure helps in maintaining diversity over the solution space in a dynamic problem. % a PSO using a dynamic and hierarchical neighborhood structure to handle dynamic optimization problems. They demonstrated that such a structure is useful for maintaining some particle diversity in a dynamic environment.

 Other forms of cooperation were also investigated in \cite{facotrsSwarm} where the problem is divided among particles and each particle solves part of the problem. \cite{facotrsSwarm} discuss a different behavior that governs cooperation particles in the swarm system.  %were used There is various topologies used in connecting a swarm in PSO. The most used is the totally connected, star and ring topology.

%\subsection*{PSO vs. Genetic Algorithms}  
% compare   between genetic programming and algorihtms.

\subsection{Drawbacks}

The main drawback of the particle swarm algorithm is the premature convergence problem. This happens when global best cannot be improved after a few swarm iterations. This is because the use of global best makes PSO converge to one place which is between local best and global best.  Another reason is because of the high amount of information between particle makes the solution less diverse which limit the explored solution space. This problem can be solved by using various methods; one of these methods is choosing another topology rather than the global best topology.  Another solution is using adaptive inertia $w$ (Section \ref{sec:ParticleSwarmAlgorithm}). Other methods of solving problem were introduced as improvement to the PSO algorithm. Another problem of PSO is that performance is dependent on the problem and the PSO parameter selected for this specific problem. This means that any change in the PSO parameter change the efficiency of PSO solution. The solution to this problem is proposed by using self adaptive parameters.

\subsection{Improvements to PSO}

There was various improvement used to improve the PSO algorithm. They mostly try to solve the problems premature converge problem and increase the diversity of the particle population. Early work on dynamic optimization was introduced by Eberhard and shi \cite{paramterSelection}, where they added the inertia weight version of PSO to find the optimum of 3 dimensional function which change every 100 iteration. Others used constrained PSO, where the general PSO Equation \ref{eq:Swarm} is constrained. Guaranteed Convergence PSO is introduced in \cite{GCPSO} where a system is guaranteed to converge to at least a local optimal solution. A repetition of this algorithm is used to insure finding of an optimal global solution.

Other hybrid system that combines PSO with other techniques was introduced. Bayesian network and probability estimation were used to move particle to better solutions. Fuzzy rules were used to improve the general PSO algorithms in \cite{shapeFuzzy}.

\subsection{Particle Swarm Applications}
\label{sec:SwarmApp}

 Since 1995, when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced Particle Swarm Optimization Algorithm (PSO) the algorithm had been used in every discipline. The algorithm was used to train Neural network in \cite{PSOandNeuarl} and proved to perform better than GA algorithm. In Image processing discipline PSO was used an optimization algorithm in \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}. PSO algorithm was used in \cite{PSOIMageClassification} to classify pixel to simplify the object segmentation process. Other system used in shape representation like \cite{shapeFuzzy}. Yin \cite{PolygonApproximationPSO} used Discrete PSO to optimally convert digital curves into polygons.

\section{Discrete Particle Swarm Algorithm}
\label{sec:ParticleSwarmAlgorithm}

  %It simulate the  % which has the best or  select a set of parameters that satisfy all constrains

\subsection{Discrete Particle Swarm Equations(DPSO)}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
where $p_{ij}$ represent the $jth$ dimension in the $ith$ particle and $v_{ij}$ is the velocity of the $jth$ dimension in the $ith$ particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} W  + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} )
\label{eq:Swarm}
\end{equation}
 where $W$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation, $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.
 After each iteration the global best $g_{best}$ particle and the agent local best $l_{best}$ particle are evaluated based on the maximum fitness functions of all particles in the solution space.

 The velocity is bounded to maximum of $V_{max}$ using Equation\ref{eq:vmax} to prevent exploding which leads to premature convergence. The solution is found after achieving a specific number of iteration or after an error threshold is achieved.
Equation \ref{eq:descrite} is used to change the general swarm algorithm into binary particle (\textit{Discrete Particle Swarm Algorithm DPSO}) which handles particle values of either $0$ or $1$ \cite{PSODisceret}. 
 \begin{equation}
   P(i)\Leftarrow 
\{
\begin{array}{c} 
1 \quad \quad if\quad r_{3}>p_{i}  \\

0 \quad \quad if\quad r_{3}<p_{i} 
\label{eq:descrite}
\end{array}\}
\end{equation} where $p_{ij}$ is the numerical values of the particle and $r_{3}$ is a random variable. 
  \begin{equation}
V_{ij}  = \frac{{V_{ij} }}{{2V_{\max } }} + 0.5
\label{eq:vmax}
\end{equation} where $V_{\max }$ is the velocity threshold.

%\subsection{The Algorithm}
%%\label{sec:ParticleSwarmAlgorithm} 
%\begin{verbatim}
% Initialize 
%    Generate N particles with M dimensions (P_i...P_N) 
%    Generate velocities for each particle for M dimension randomly. 
% Repeat till Maximum number of iteration OR Fitness 
%         For each particle P_i do the following 
%                Compute fitness  
%                Determine Lbest  
%                Check if need to change Gbest. 
%                Compute the Velocity for each dimension using eq. 
%                Check if maximum Velocity reach and limit it 
%                update particle values 
%          end . 
%        Improve all particle positions 
%  end 
%  
%  Final solution reaches in Gbest.       
%        
%        
%\end{verbatim}
